---
title: "Differential effects of advertisement strategies on vaccine uptake: data analysis"
author: "Aspen Han"
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk:
    out.width: "75%"
execute:
  echo: false
format:
  pdf:
    pdf-engine: xelatex
    keep-tex: true
    documentmode: man
    mainfont: Times New Roman
    sansfont: Garamond
    mathfont: Cambria Math
    fontsize: "12"
    linestretch: 2
    top-level-division: section
    floatsintext: true
    fig-width: 6
    fig-asp: 0.618
    fig-align: center
    include-in-header:
      text: |
        \usepackage{sectsty}
        \sectionfont{\centering}
        \subsectionfont{\raggedright}
        \usepackage{indentfirst}
        \setlength{\parindent}{3em}
        \usepackage{amsmath}
        \usepackage{amssymb}
        \usepackage{enumitem}
        \setlist[itemize]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
        \setlist[enumerate]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
        \usepackage{caption}
---

```{r}
#| label: setup
#| include: false

#Load packages

## Data reading and writing, transformation, and visualisation
library(tidyverse)
library(broom) #turns function output into tidy tibbles

## Statistical analysis
library(DescTools) #descriptive stats
library(car) #basic regression analysis
library(sandwich) #estimate robust errors
library(estimatr) #run robust and other more complex regressions and analysis
library(AER) #run IV regressions
library(nnet) #run multinomial logistic regressions
library(fixest) #run fixed effects regressions
library(lmtest) #run diagnostic tests on linear models

## Tables
library(texreg) #make regression tables
library(gtsummary) #make summary stats tables
library(gt) #make general tables

```

```{r}
#| label: default-plot-themes
#| include: false

# Set default theme for ggplot; serif is times new roman

def_theme <- theme(plot.title = element_text(family = "serif", size = 13, hjust = 0.5),
                   axis.title = element_text(family = "serif", size = 12),
                   axis.text = element_text(family = "serif", size = 11),
                   strip.text.x = element_text(family = "serif", size = 12),
                   strip.text.y = element_text(family = "serif", size = 12, angle = 0),
                   legend.title = element_text(family = "serif", size = 12),
                   legend.text = element_text(family = "serif", size = 11))

```

```{r}
#| label: read-in-data
#| include: false

treatment <- read_csv("treatment-data.csv")

endline <- read_csv("endline-data.csv")

```

```{r}
#| label: convert-data-types

treatment <- treatment %>%
    mutate(gender = factor(gender), race = factor(race),
         edu = factor(edu, levels = c("below high school", "high school", "some college", "bachelor's or above")),
         income_bracket = factor(income_bracket, levels = c("<25k", "25k-<50k", "50k-<75k", "75k-<120k", "120k-<200k", ">=200k")),
         state = factor(state),
         age_group = factor(age_group, levels = c("<30", "30-44", "45-64", "65+")),
         treatment = factor(treatment, levels = c("pathos", "logos", "control")))
         
endline <- endline %>%
  mutate(gender = factor(gender), race = factor(race),
         edu = factor(edu, levels = c("below high school", "high school", "some college", "bachelor's or above")),
         income_bracket = factor(income_bracket, levels = c("<25k", "25k-<50k", "50k-<75k", "75k-<120k", "120k-<200k", ">=200k")),
         state = factor(state),
         age_group = factor(age_group, levels = c("<30", "30-44", "45-64", "65+")),
         treatment = factor(treatment, levels = c("pathos", "logos", "control")))

```

Following the simulation of the field experiment, I turn to analyze its data. Foremost, I check whether attrition has eroded the internal validity of the study, before proceeding to check for average and differential effects of Facebook ad campaigns on Covid-19 vaccine uptake.

## Covariate balance post-attrition

Using covariates[^1] as a proxy for potential outcomes, I first check whether they are balanced between treatment groups in the endline sample, as presented in @tbl-cov-balance-endline. This checks whether we can recover causal ATE estimates among the respondents, and indeed the table shows no significant differences in any of the covariates between treatment groups. Note if there were significant differences found at the individual covariate level, one would still have to correct for multiple hypotheses testing, or just a joint test of covariates to ascertain actual differences.

I also check for covariate balance between the baseline and endline sample, which helps to inform whether we can recover ATE for the whole sample. @tbl-cov-balance-full shows that there is a sole significant difference in initial willingness to take the updated 2024-2025 Covid-19 vaccine (p-value = 0.043) although this will likely not hold after correcting for multiple hypotheses testing or using a joint test.

Hence, it is likely that causal treatment effects can be consistently estimated, though this was somewhat by construction since I simulated the attrition to be random. In reality, treatment receipt may affect attrition if for some reason people exposed to the ad campaigns were more or less likely to return for the endline survey (e.g. they are more likely as they felt more engaged by the increased ad exposure) which would reintroduce selection bias effects. Running the above tests are important to establish continued internal validity of longitudinal studies like this which are susceptible to attrition.

[^1]: States have been omitted from this and subsequent analysis and it introduces too much dimensionality and restrictions on degrees of freedom.

```{r}
#| label: tbl-cov-balance-endline
#| tbl-cap: "Covariate balance across treatment groups in the endline sample"

# Create summary statistics of covariates

cov_sumstats_endline <- tbl_summary(endline, #dataframe to summarise
                                    missing = "no", #do not state number of observations with missing values
                                    by = treatment, #group by treatment
                                    include = c(gender, race, age_group, edu, income_bracket, fb_usage, vax_percpt), #select variables to include
                                    statistic = list(all_continuous() ~ "{mean} ({sd})",
                                                     all_categorical() ~ "{n} ({p}%)") #specify sumstats to compute for each variable type
) %>%
  add_p() %>%
  bold_p()

cov_sumstats_endline

```

```{r}
#| label: tbl-cov-balance-full
#| tbl-cap: "Covariate balance across baseline and endline"

# create a variable for whether subjects in the full sample responded at endline

treatment$endline_response <- ifelse(treatment$identifier %in% endline$identifier, 1, 0)

cov_sumstats_full <- tbl_summary(treatment,
                                 by = endline_response,
                                 include = c(gender, race, age_group, edu, income_bracket, fb_usage, vax_percpt),
                                 statistic = list(all_continuous() ~ "{mean} ({sd})",
                                                  all_categorical() ~ "{n} ({p}%)")
) %>%
  add_p() %>%
  bold_p()
                                 
                                 
cov_sumstats_full

```

## Average treatment effects from ad exposure

I regress the new willingness to take the updated vaccine on a treatment dummy for whether the subject was exposed to an ad campaign at all (and other covariates), measuring an intention to treat (ITT) effect. I also use treatment as an instrument for ad awareness, to measure the local average treatment effect (LATE) of the ad campaign on vaccine uptake among those who were actually engaged with it.

@tbl-reg-avg-effects shows the results. Both ITT and LATE (coefficients on `ad_exposed` and `ad_aware` respectively) are significant and quite similar, with LATE being stronger as expected since people who engage more with the ads are likely to be more influenced by them. Controlling for covariates does not affect the ITT estimate much though it reduces the standard error considerably. Overall, ad campaigns are associated with around a 0.4 increase in willingness to take the updated vaccine, which corresponds to a jump up of almost half a category, which is quite substantial given the campaign only lasted one month, though it is difficult to say how effects would scale with time.

In terms of covariate effects, one notable finding is that initial vaccine perceptions show a significantly high and positive correlation with later vaccine perceptions, suggesting people who already viewed the vaccines more favourably to begin are more amenable to them in the future. Given its huge standalone effect, I considered its interplay with ad campaign exposure effects, which found a significantly negative interaction effect (coefficient of -0.22, p-value < 0.01). This means ad campaigns are not only less effective on those who already were willing to get the updated effect but may reverse their perceptions, perhaps due to annoyance or crowding out effects, so there are perhaps unintended consequences on top of limited scope of gains we have to bear in mind when deciding who to target such ads to.

```{r}
#| label: create-variables

endline <- endline %>%
  mutate(ad_exposed = ifelse(treatment == "control", 0, 1),
         ad_aware = ifelse(ad_awareness == "No", 0, 1))

```

```{r}
#| label: run-regression-average-effects

# ITT regression with only the treatment dummy as a regressor

reg_itt_basic <- lm_robust(new_vax_percpt ~ ad_exposed, endline)

# ITT regression with other covariates as regressors also

reg_itt_ext <- lm_robust(new_vax_percpt ~ ad_exposed + gender + race + age_group + edu + income_bracket + fb_usage + vax_percpt, endline)

# LATE regression
reg_late <- ivreg(formula = (new_vax_percpt ~ ad_aware + gender + race + age_group + edu + income_bracket + fb_usage + vax_percpt|ad_exposed + gender + race + age_group + edu + income_bracket + fb_usage + vax_percpt), data = endline)

reg_models_avg <- list(reg_itt_basic, reg_itt_ext, reg_late)

```

```{r}
#| label: tabulate-regression-average-effects
#| include: false

reg_table_avg <- texreg(reg_models_avg, #regression outputs to use
                        stars = c(0.01, 0.05, 0.1), #significance indicators
                        include.ci = FALSE, #use standard errors for estimates
                        custom.model.names = c("ITT: No controls", "ITT: With controls", "LATE"),
                    #name models
                        custom.header = list("New willingness for vaccine uptake" = 1:3),
                    #label dependent variable above model names
                        include.rsquared = FALSE,
                        include.rmse = FALSE,
                        caption = NULL)

reg_table_avg

```

::: {#tbl-reg-avg-effects}

\begin{table}
\begin{center}
\begin{tabular}{l c c c}
\hline
 & \multicolumn{3}{c}{New willingness for vaccine uptake} \\
\cline{2-4}
 & ITT: No controls & ITT: With controls & LATE \\
\hline
(Intercept)                 & $2.75^{***}$ & $0.34^{***}$  & $0.41^{***}$ \\
                            & $(0.04)$     & $(0.04)$      & $(0.05)$     \\
ad\_exposed                 & $0.37^{***}$ & $0.40^{***}$  &              \\
                            & $(0.05)$     & $(0.01)$      &              \\
gendermale                  &              & $0.04^{***}$  & $0.03^{**}$  \\
                            &              & $(0.01)$      & $(0.01)$     \\
raceblack                   &              & $0.02$        & $0.01$       \\
                            &              & $(0.03)$      & $(0.03)$     \\
racehispanic                &              & $-0.01$       & $-0.00$      \\
                            &              & $(0.03)$      & $(0.03)$     \\
raceother                   &              & $-0.05$       & $-0.05$      \\
                            &              & $(0.05)$      & $(0.06)$     \\
racewhite                   &              & $-0.01$       & $-0.01$      \\
                            &              & $(0.03)$      & $(0.03)$     \\
age\_group30-44             &              & $-0.03$       & $-0.02$      \\
                            &              & $(0.02)$      & $(0.02)$     \\
age\_group45-64             &              & $-0.01$       & $-0.00$      \\
                            &              & $(0.02)$      & $(0.02)$     \\
age\_group65+               &              & $-0.05^{***}$ & $-0.05^{**}$ \\
                            &              & $(0.02)$      & $(0.02)$     \\
eduhigh school              &              & $-0.03$       & $-0.03$      \\
                            &              & $(0.03)$      & $(0.03)$     \\
edusome college             &              & $0.00$        & $0.01$       \\
                            &              & $(0.03)$      & $(0.03)$     \\
edubachelor's or above      &              & $-0.04$       & $-0.03$      \\
                            &              & $(0.03)$      & $(0.04)$     \\
income\_bracket25k-$<$50k   &              & $0.03$        & $0.04$       \\
                            &              & $(0.03)$      & $(0.03)$     \\
income\_bracket50k-$<$75k   &              & $0.04$        & $0.05$       \\
                            &              & $(0.03)$      & $(0.03)$     \\
income\_bracket75k-$<$120k  &              & $0.03$        & $0.04$       \\
                            &              & $(0.04)$      & $(0.04)$     \\
income\_bracket120k-$<$200k &              & $0.04$        & $0.06$       \\
                            &              & $(0.04)$      & $(0.05)$     \\
income\_bracket$>$=200k     &              & $0.09$        & $0.09$       \\
                            &              & $(0.07)$      & $(0.08)$     \\
fb\_usage                   &              & $0.01^{***}$  & $-0.01^{**}$ \\
                            &              & $(0.00)$      & $(0.00)$     \\
vax\_percpt                 &              & $0.86^{***}$  & $0.86^{***}$ \\
                            &              & $(0.00)$      & $(0.00)$     \\
ad\_aware                   &              &               & $0.43^{***}$ \\
                            &              &               & $(0.02)$     \\
\hline
Adj. R$^2$                  & $0.02$       & $0.90$        & $0.90$       \\
Num. obs.                   & $4500$       & $4500$        & $4500$       \\
\hline
\multicolumn{4}{l}{\scriptsize{$^{***}p<0.01$; $^{**}p<0.05$; $^{*}p<0.1$}}
\end{tabular}
\label{table:coefficients}
\end{center}
\end{table}

Average effects of Facebook ad campaigns on vaccine uptake

:::

```{r}
#| label: interaction-effects
#| include: false

reg_itt_int <- lm_robust(new_vax_percpt ~ ad_exposed + gender + race + age_group + edu + income_bracket + fb_usage + vax_percpt + ad_exposed*vax_percpt, endline)

reg_itt_int

```


## Differential effects of ad campaign strategies

Finally, I am interested in how logos and pathos used in ads differ in their efficacy of increasing willingness for vaccine uptake. This time, I regress the same outcome variable but on the treatment groups, with pathos as the baseline (omitted regressor). @tbl-reg-avg-effects shows that ad campaigns appealing to reason are more effective at encouraging vaccine uptake than those appealing to emotions, being a third more likely to induce people to move up a category in their likelihood of getting an updated vaccine shot. It would be interesting to understand what features of the context can contribute to these differences.

```{r}
#| label: run-regression-differential-effects

reg_diff_basic <- lm_robust(new_vax_percpt ~ treatment, endline)

reg_diff_ext <- lm_robust(new_vax_percpt ~ treatment + gender + race + age_group + edu + income_bracket + fb_usage + vax_percpt, endline)

reg_models_diff <- list(reg_diff_basic, reg_diff_ext)

```

```{r}
#| label: tabulate-regression-differential-effects
#| include: false

reg_table_diff <- texreg(reg_models_diff, #regression outputs to use
                         stars = c(0.01, 0.05, 0.1), #significance indicators
                         include.ci = FALSE, #use standard errors for estimates
                         custom.model.names = c("No controls", "With controls"),
                    #name models
                         custom.header = list("New willingness for vaccine uptake" = 1:2),
                    #label dependent variable above model names
                         include.rsquared = FALSE,
                         include.rmse = FALSE,
                         caption = NULL)

reg_table_diff

```

::: {#tbl-reg-avg-effects}

\begin{table}
\begin{center}
\begin{tabular}{l c c}
\hline
 & \multicolumn{2}{c}{New willingness for vaccine uptake} \\
\cline{2-3}
 & No controls & With controls \\
\hline
(Intercept)                 & $2.90^{***}$  & $0.56^{***}$  \\
                            & $(0.04)$      & $(0.04)$      \\
treatmentlogos              & $0.45^{***}$  & $0.36^{***}$  \\
                            & $(0.05)$      & $(0.02)$      \\
treatmentcontrol            & $-0.15^{***}$ & $-0.21^{***}$ \\
                            & $(0.06)$      & $(0.01)$      \\
gendermale                  &               & $0.04^{***}$  \\
                            &               & $(0.01)$      \\
raceblack                   &               & $0.02$        \\
                            &               & $(0.03)$      \\
racehispanic                &               & $-0.01$       \\
                            &               & $(0.03)$      \\
raceother                   &               & $-0.06$       \\
                            &               & $(0.05)$      \\
racewhite                   &               & $-0.01$       \\
                            &               & $(0.03)$      \\
age\_group30-44             &               & $-0.02$       \\
                            &               & $(0.02)$      \\
age\_group45-64             &               & $-0.01$       \\
                            &               & $(0.02)$      \\
age\_group65+               &               & $-0.05^{**}$  \\
                            &               & $(0.02)$      \\
eduhigh school              &               & $-0.02$       \\
                            &               & $(0.03)$      \\
edusome college             &               & $0.01$        \\
                            &               & $(0.03)$      \\
edubachelor's or above      &               & $-0.03$       \\
                            &               & $(0.03)$      \\
income\_bracket25k-$<$50k   &               & $0.03$        \\
                            &               & $(0.03)$      \\
income\_bracket50k-$<$75k   &               & $0.03$        \\
                            &               & $(0.03)$      \\
income\_bracket75k-$<$120k  &               & $0.03$        \\
                            &               & $(0.04)$      \\
income\_bracket120k-$<$200k &               & $0.03$        \\
                            &               & $(0.04)$      \\
income\_bracket$>$=200k     &               & $0.06$        \\
                            &               & $(0.07)$      \\
fb\_usage                   &               & $0.01^{***}$  \\
                            &               & $(0.00)$      \\
vax\_percpt                 &               & $0.86^{***}$  \\
                            &               & $(0.00)$      \\
\hline
Adj. R$^2$                  & $0.03$        & $0.91$        \\
Num. obs.                   & $4500$        & $4500$        \\
\hline
\multicolumn{3}{l}{\scriptsize{$^{***}p<0.01$; $^{**}p<0.05$; $^{*}p<0.1$}}
\end{tabular}
\caption{}
\label{table:coefficients}
\end{center}
\end{table}

Differential effects of Facebook ad campaigns on vaccine uptake

:::